---
title: "Engineering Rolling Average Variable"
author: "Shruti Gopalswamy"
output: html_notebook
---

# Loading neccessary libraries
```{r}
library(purrr)
library(dplyr)
library(data.table)
library(tidyverse)
library(tidyr)
library(stringr)
```

# Importing the Data
```{r}
games <-  read.csv("https://raw.githubusercontent.com/mattymo18/STOR-538-Project2-2021/master/Source-Data/games.csv")

teams <- read.csv("https://raw.githubusercontent.com/mattymo18/STOR-538-Project2-2021/master/Source-Data/teams.csv")

game_details_cleaned <- read.csv("https://raw.githubusercontent.com/john-lickteig/playoff2/master/game_details_cleaned.csv")

games_cleaned <- read.csv("https://raw.githubusercontent.com/john-lickteig/playoff2/master/games_cleaned.csv")

all_game_data <- read.csv("https://raw.githubusercontent.com/john-lickteig/playoff2/master/all_game_data.csv")
```


# Created a list of tables, where each table represents one team's home games. Then, kept only columns for GAME_DATE_EST, HOME_TEAM_ID, and PTS_home in each table. 
```{r}
HomeGamesByTeamList <- split(games_cleaned, as.factor(games_cleaned$HOME_TEAM_ID))

HomeGamesByTeamList <- lapply(HomeGamesByTeamList, function(x) x%>% select(GAME_DATE_EST, HOME_TEAM_ID, PTS_home))
```

# Since each table in the HomeGamesByTeamList is already sorted in order of descending GAME_DATE_EST, just kept the top 5 rows, representing the 5 most recent game played by that team. Then calculated the average points scored by the home team in the last 5 games.
```{r}
ListofHomeGamesByTeam_PtsOnly <- HomeGamesByTeamList %>%
  lapply(function(x) x%>% select(PTS_home)) %>% 
  lapply(function(x) x%>% slice(1:5)) 

AvgPtsForRecent5HomeGamesByTeam <- as.data.frame(do.call(rbind,lapply(ListofHomeGamesByTeam_PtsOnly, function(x) { colSums(x) / 5 }))) %>%
  setDT(keep.rownames = "TEAM_ID")

  colnames(AvgPtsForRecent5HomeGamesByTeam)[2] <- "LAST_5_AVG_PTS_home"
```



# Created a list of tables, where each table represents one team's away games. Then, kept only columns for GAME_DATE_EST, VISITOR_TEAM_ID, and PTS_away in each table. 
```{r}
AwayGamesByTeamList <- split(games_cleaned, as.factor(games_cleaned$VISITOR_TEAM_ID))

AwayGamesByTeamList <- lapply(AwayGamesByTeamList, function(x) x%>% select(GAME_DATE_EST, VISITOR_TEAM_ID, PTS_away))
```

# Since each table in the AwayGamesByTeamList is already sorted in order of descending GAME_DATE_EST, just kept the top 5 rows, representing the 5 most recent away games played by that team. Then calculated the average points scored by the away team in the last 5 games.

```{r}
ListofAwayGamesByTeam_PtsOnly <- AwayGamesByTeamList %>%
  lapply(function(x) x%>% select(PTS_away)) %>% 
  lapply(function(x) x%>% slice(1:5)) 

AvgPtsForRecent5AwayGamesByTeam <- as.data.frame(do.call(rbind,lapply(ListofAwayGamesByTeam_PtsOnly, function(x) { colSums(x) / 5 }))) %>%
  setDT(keep.rownames = "TEAM_ID")

  colnames(AvgPtsForRecent5AwayGamesByTeam)[2] <- "LAST_5_AVG_PTS_away"
```


# Creating Home and Away Team Variables
```{r}
team_names <- teams %>%
            select(TEAM_ID,CITY,NICKNAME) %>%
            unite(NAME,CITY,NICKNAME,sep=" ")
team_names$TEAM_ID <- as.character(team_names$TEAM_ID)
head(team_names)
```

# Merging both AVG_PTS_home and AVG_PTS_away into one data frame and exporting the csv
```{r}
AvgPtsForRecent5GamesByTeam = merge(x = AvgPtsForRecent5HomeGamesByTeam, y = AvgPtsForRecent5AwayGamesByTeam, by = "TEAM_ID", all.x = TRUE)

 AvgPtsForRecent5GamesByTeam = merge(x = team_names, y = AvgPtsForRecent5GamesByTeam, by = "TEAM_ID", all.x = TRUE) 
```



