---
title: "Playoff Round 2"
output: html_notebook
---

In the following notebook, we use data from NBA games to create models that predict point spread, total points scored, and offensive rebounds. 

```{r}
library(tidyverse)
library(tidyr)
library(stringr)
library(ggplot2)
library(lubridate)
library(dplyr)
```

# Importing the Data
```{r}
games <-  read.csv("https://raw.githubusercontent.com/mattymo18/STOR-538-Project2-2021/master/Source-Data/games.csv")
game_details <- read.csv("https://raw.githubusercontent.com/mattymo18/STOR-538-Project2-2021/master/Source-Data/games_details.csv")
teams <- read.csv("https://raw.githubusercontent.com/mattymo18/STOR-538-Project2-2021/master/Source-Data/teams.csv")
```

# Previewing Data
```{r}
head(games)
head(game_details)
head(teams)
```


*Part 1 - Cleaning the games dataset*

```{r}
head(games)
```

# Isolating NA columns in games
```{r}
games_na <- games[rowSums(is.na(games)) > 0, ]
head(games_na, 99)
```
99 observations are missing all in-game statistics. We will drop these rows from the dataset.

```{r}
games$GAME_DATE_EST <- as.Date(games$GAME_DATE_EST)
games_final <- games %>%
  drop_na() %>%
  filter(GAME_DATE_EST >= "2014-10-28")
head(games_final)
write.csv(games_final, 'games_cleaned.csv')
```


*Part 2 - Cleaning the game_details dataset*

```{r}
head(game_details, 300)
```

# Adding dates to player performances
```{r}
game_details_dates <- game_details %>%
  left_join(games_final, by="GAME_ID") %>%
  select(GAME_DATE_EST, colnames(game_details))
head(game_details_dates)
```

# Removing observations from games before 2014

```{r}
game_details_dates$GAME_DATE_EST <- as.Date(game_details_dates$GAME_DATE_EST)
head(game_details_dates, 600)
```


```{r}
game_details_new <- game_details_dates %>%
  filter(GAME_DATE_EST >= "2014-10-28")
head(game_details_new)
```

```{r}
game_details_final <- game_details_new %>%
  filter(COMMENT == "")
head(game_details_final)
write.csv(game_details_final, 'game_details_cleaned.csv')
```

*Part 3 - Offensive Rebounding Cleaning*

# Condense Game Data
```{r}
games_data <- games_final %>%
            select(GAME_DATE_EST,GAME_ID,HOME_TEAM_ID,VISITOR_TEAM_ID,PTS_home,PTS_away,HOME_TEAM_WINS) %>%
            mutate(Spread=PTS_home-PTS_away,Total=PTS_home+PTS_away)
head(games_data)
```

# Aggregate all player stats from each game and assign to team's respective GAME_ID
```{r}
sum_stats = game_details_final %>%
          select (TEAM_ABBREVIATION,GAME_ID,TEAM_ID,FGM,FGA,FG3M,FG3A,FTM,FTA,OREB,DREB,AST,STL,BLK,TO,PF,PLUS_MINUS) %>%
          group_by(TEAM_ABBREVIATION,GAME_ID,TEAM_ID) %>%
          summarize(OREB=sum(OREB,na.rm=T),
                    DREB=sum(DREB,na.rm=T),
                    FGM=sum(FGM,na.rm=T),
                    FGA=sum(FGA,na.rm=T),
                    FG3M=sum(FG3M,na.rm=T),
                    FG3A=sum(FG3A,na.rm=T),
                    FTM=sum(FTM,na.rm=T),
                    FTA=sum(FTA,na.rm=T),
                    AST=sum(AST,na.rm=T),
                    STL=sum(STL,na.rm=T),
                    BLK=sum(BLK,na.rm=T),
                    TO=sum(TO,na.rm=T),
                    PF=sum(PF,na.rm=T),
                    PLUS_MINUS=sum(PLUS_MINUS,na.rm=T)) %>%
          ungroup()
head(sum_stats)
```


# Merge All Aggregated Team Stats into Game Data
```{r}
final_game_data = left_join(games_data,select(OREB,-TEAM_ABBREVIATION),by=c("GAME_ID","HOME_TEAM_ID"="TEAM_ID")) %>%
                    rename(OREB_home=OREB) %>%
                    left_join(select(OREB,-TEAM_ABBREVIATION),by=c("GAME_ID","VISITOR_TEAM_ID"="TEAM_ID")) %>%
                    rename(OREB_away=OREB) %>%
                    mutate(OREB=OREB_home+OREB_away)

head(final_game_data)
```

#Join game data with each team and point spread
```{r}
all_box_stats <- games_data %>%
  left_join(sum_stats,select(OREB,DREB,FGM,FGA,FG3M,FG3A,FTM,FTA,AST,STL,BLK,TO,PF,-TEAM_ABBREVIATION),by=c("GAME_ID","HOME_TEAM_ID"="TEAM_ID")) %>%
  rename(OREB_home=OREB,
         DREB_home=DREB,
         FGM_home=FGM,
         FGA_home=FGA,
         FG3M_home=FG3M,
         FG3A_home=FG3A,
         FTM_home=FTM,
         FTA_home=FTA,
         AST_home=AST,
         STL_home=STL,
         BLK_home=BLK,
         TO_home=TO,
         PF_home=PF,
         PM_home=PLUS_MINUS) %>%
  left_join(sum_stats,select(OREB,DREB,FGM,FGA,FG3M,FG3A,FTM,FTA,AST,STL,BLK,TO,PF,-TEAM_ABBREVIATION),by=c("GAME_ID","VISITOR_TEAM_ID"="TEAM_ID")) %>%
  rename(OREB_away=OREB,
         DREB_away=DREB,
         FGM_away=FGM,
         FGA_away=FGA,
         FG3M_away=FG3M,
         FG3A_away=FG3A,
         FTM_away=FTM,
         FTA_away=FTA,
         AST_away=AST,
         STL_away=STL,
         BLK_away=BLK,
         TO_away=TO,
         PF_away=PF,
         PM_away=PLUS_MINUS) %>%
  mutate(OREB=OREB_home+OREB_away,
         DREB=DREB_home+DREB_away,
         FGM=FGM_home+FGM_away,
         FGA=FGA_home+FGA_away,
         FG3M=FG3M_home+FG3M_away,
         FG3A=FG3A_home+FG3A_away,
         FTM=FTM_home+FTM_away,
         FTA=FTA_home+FTA_away,
         AST=AST_home+AST_away,
         STL=STL_home+STL_away,
         BLK=BLK_home+BLK_away,
         TO=TO_home+TO_away,
         PF=PF_home+PF_away) %>%
  subset(select=-c(TEAM_ABBREVIATION.x,TEAM_ABBREVIATION.y))

head(all_box_stats)
```

#Reorder columns so home, away, and total metrics are all next to one another
```{r}
all_box_stats <- all_box_stats[ ,c("GAME_DATE_EST", "GAME_ID", "HOME_TEAM_ID", "VISITOR_TEAM_ID", "PTS_home", "PTS_away", "Spread", "Total", "OREB_home", "OREB_away", "OREB", "DREB_home", "DREB_away", "DREB", "FGM_home", "FGM_away", "FGM", "FGA_home", "FGA_away", "FGA", "FG3M_home", "FG3M_away", "FG3M", "FG3A_home", "FG3A_away", "FG3A", "FTM_home", "FTM_away", "FTM", "FTA_home", "FTA_away", "FTA", "AST_home", "AST_away", "AST", "STL_home", "STL_away", "STL", "BLK_home", "BLK_away", "BLK", "TO_home", "TO_away", "TO", "PF_home", "PF_away", "PF", "PM_home", "PM_away", "HOME_TEAM_WINS")]
head(all_box_stats)
```



# Creating Home and Away Team Variables
```{r}
team_names <- teams %>%
            select(TEAM_ID,CITY,NICKNAME) %>%
            unite(NAME,CITY,NICKNAME,sep=" ")
head(team_names)
```

# Merging Team Name into Original Data
```{r}
final_game_data = left_join(all_box_stats,team_names,by=c("HOME_TEAM_ID"="TEAM_ID")) %>%
                        rename("Home Team"=NAME) %>%
                        left_join(team_names,by=c("VISITOR_TEAM_ID"="TEAM_ID")) %>%
                        rename("Away Team"=NAME) %>%
                        select(GAME_DATE_EST,"Home Team","Away Team",everything()) %>%
                        select(-GAME_ID,-HOME_TEAM_ID,-VISITOR_TEAM_ID)
head(final_game_data)
```

```{r}
view(final_game_data)
```


```{r}
write.csv(final_game_data, 'final_game_data.csv')
```


```{r}
rolling_avg <-read.csv("https://raw.githubusercontent.com/john-lickteig/playoff2/main/final_game_data_with_rolling_avg.csv")
head(rolling_avg)
```

# Selecting needed variables for season_level stats
```{r}
season_level_avg <-  all_box_stats[, c("GAME_DATE_EST", "GAME_ID", "HOME_TEAM_ID", "VISITOR_TEAM_ID", "Spread", "Total", "OREB", "PTS_home", "PTS_away", "OREB_home", "OREB_away", "DREB_home", "DREB_away", "FGM_home", "FGM_away", "FGA_home", "FGA_away", "FG3M_home", "FG3M_away", "FG3M", "FG3A_home", "FG3A_away", "FTM_home", "FTM_away", "FTA_home", "FTA_away",  "AST_home", "AST_away", "STL_home", "STL_away", "BLK_home", "BLK_away", "TO_home", "TO_away", "PF_home", "PF_away", "PM_home", "PM_away", "HOME_TEAM_WINS")]

head(season_level_avg)
```

```{r}
season_level_avg = left_join(season_level_avg,team_names,by=c("HOME_TEAM_ID"="TEAM_ID")) %>%
                        rename(Home.Team=NAME) %>%
                        left_join(team_names,by=c("VISITOR_TEAM_ID"="TEAM_ID")) %>%
                        rename(Away.Team=NAME)

head(season_level_avg)
```

```{r}
season_avg_ordered <- season_level_avg %>%
  select(GAME_DATE_EST,Home.Team,Away.Team,everything())

head(season_avg_ordered)
```



# Find home team season-level averages for 2014-2019
```{r}
season_level_avg_2014 <- season_avg_ordered %>%
  filter(GAME_DATE_EST <= "2015-4-15") %>%
  group_by(HOME_TEAM_ID) %>%
  summarize(GAME_DATE_EST = GAME_DATE_EST,
            GAME_ID = GAME_ID,
            PTS_home = PTS_home,
            games_each = n(),
            PTS_home_avg = sum(PTS_home)/games_each,
            OREB_home_avg = sum(OREB_home)/games_each,
            DREB_home_avg = sum(DREB_home)/games_each,
            FGM_home_avg = sum(FGM_home)/games_each,
            FGA_home_avg = sum(FGA_home)/games_each,
            FG3M_home_avg = sum(FG3M_home)/games_each,
            FG3A_home_avg = sum(FG3A_home)/games_each,
            FTM_home_avg = sum(FTM_home)/games_each,
            FTA_home_avg = sum(FTA_home)/games_each,
            AST_home_avg = sum(AST_home)/games_each,
            STL_home_avg = sum(STL_home)/games_each,
            BLK_home_avg = sum(BLK_home)/games_each,
            TO_home_avg = sum(TO_home)/games_each,
            PF_home_avg = sum(PF_home)/games_each,
            PM_home_avg = sum(PM_home)/games_each,
            PTS_home_against_avg = sum(PTS_away)/games_each,
            OREB_home_against_avg = sum(OREB_away)/games_each,
            DREB_home_against_avg = sum(DREB_away)/games_each,
            FGM_home_against_avg = sum(FGM_away)/games_each,
            FGA_home_against_avg = sum(FGA_away)/games_each,
            FG3M_home_against_avg = sum(FG3M_away)/games_each,
            FG3A_home_against_avg = sum(FG3A_away)/games_each,
            FTM_home_against_avg = sum(FTM_away)/games_each,
            FTA_home_against_avg = sum(FTA_away)/games_each,
            AST_home_against_avg = sum(AST_away)/games_each,
            STL_home_against_avg = sum(STL_away)/games_each,
            BLK_home_against_avg = sum(BLK_away)/games_each,
            TO_home_against_avg = sum(TO_away)/games_each,
            PF_home_against_avg = sum(PF_away)/games_each,
            PM_home_against_avg = sum(PM_away)/games_each)


head(season_level_avg_2014, 40)

```

```{r}
season_level_avg_2015 <- season_avg_ordered %>%
  filter(GAME_DATE_EST >= "2015-10-27" & GAME_DATE_EST <= "2016-4-13") %>%
  group_by(HOME_TEAM_ID) %>%
  summarise(GAME_DATE_EST = GAME_DATE_EST,
            GAME_ID = GAME_ID,
            PTS_home = PTS_home,
            games_each = n(),
            PTS_home_avg = sum(PTS_home)/games_each,
            OREB_home_avg = sum(OREB_home)/games_each,
            DREB_home_avg = sum(DREB_home)/games_each,
            FGM_home_avg = sum(FGM_home)/games_each,
            FGA_home_avg = sum(FGA_home)/games_each,
            FG3M_home_avg = sum(FG3M_home)/games_each,
            FG3A_home_avg = sum(FG3A_home)/games_each,
            FTM_home_avg = sum(FTM_home)/games_each,
            FTA_home_avg = sum(FTA_home)/games_each,
            AST_home_avg = sum(AST_home)/games_each,
            STL_home_avg = sum(STL_home)/games_each,
            BLK_home_avg = sum(BLK_home)/games_each,
            TO_home_avg = sum(TO_home)/games_each,
            PF_home_avg = sum(PF_home)/games_each,
            PM_home_avg = sum(PM_home)/games_each,
            PTS_home_against_avg = sum(PTS_away)/games_each,
            OREB_home_against_avg = sum(OREB_away)/games_each,
            DREB_home_against_avg = sum(DREB_away)/games_each,
            FGM_home_against_avg = sum(FGM_away)/games_each,
            FGA_home_against_avg = sum(FGA_away)/games_each,
            FG3M_home_against_avg = sum(FG3M_away)/games_each,
            FG3A_home_against_avg = sum(FG3A_away)/games_each,
            FTM_home_against_avg = sum(FTM_away)/games_each,
            FTA_home_against_avg = sum(FTA_away)/games_each,
            AST_home_against_avg = sum(AST_away)/games_each,
            STL_home_against_avg = sum(STL_away)/games_each,
            BLK_home_against_avg = sum(BLK_away)/games_each,
            TO_home_against_avg = sum(TO_away)/games_each,
            PF_home_against_avg = sum(PF_away)/games_each,
            PM_home_against_avg = sum(PM_away)/games_each)

head(season_level_avg_2015, 30)
```

```{r}
season_level_avg_2016 <- season_avg_ordered %>%
  filter(GAME_DATE_EST >= "2016-10-25" & GAME_DATE_EST <= "2017-4-13") %>%
  group_by(HOME_TEAM_ID) %>%
  summarise(GAME_DATE_EST = GAME_DATE_EST,
            GAME_ID = GAME_ID,
            PTS_home = PTS_home,
            games_each = n(),
            PTS_home_avg = sum(PTS_home)/games_each,
            OREB_home_avg = sum(OREB_home)/games_each,
            DREB_home_avg = sum(DREB_home)/games_each,
            FGM_home_avg = sum(FGM_home)/games_each,
            FGA_home_avg = sum(FGA_home)/games_each,
            FG3M_home_avg = sum(FG3M_home)/games_each,
            FG3A_home_avg = sum(FG3A_home)/games_each,
            FTM_home_avg = sum(FTM_home)/games_each,
            FTA_home_avg = sum(FTA_home)/games_each,
            AST_home_avg = sum(AST_home)/games_each,
            STL_home_avg = sum(STL_home)/games_each,
            BLK_home_avg = sum(BLK_home)/games_each,
            TO_home_avg = sum(TO_home)/games_each,
            PF_home_avg = sum(PF_home)/games_each,
            PM_home_avg = sum(PM_home)/games_each,
            PTS_home_against_avg = sum(PTS_away)/games_each,
            OREB_home_against_avg = sum(OREB_away)/games_each,
            DREB_home_against_avg = sum(DREB_away)/games_each,
            FGM_home_against_avg = sum(FGM_away)/games_each,
            FGA_home_against_avg = sum(FGA_away)/games_each,
            FG3M_home_against_avg = sum(FG3M_away)/games_each,
            FG3A_home_against_avg = sum(FG3A_away)/games_each,
            FTM_home_against_avg = sum(FTM_away)/games_each,
            FTA_home_against_avg = sum(FTA_away)/games_each,
            AST_home_against_avg = sum(AST_away)/games_each,
            STL_home_against_avg = sum(STL_away)/games_each,
            BLK_home_against_avg = sum(BLK_away)/games_each,
            TO_home_against_avg = sum(TO_away)/games_each,
            PF_home_against_avg = sum(PF_away)/games_each,
            PM_home_against_avg = sum(PM_away)/games_each)

head(season_level_avg_2016, 30)
```

```{r}
season_level_avg_2017 <- season_avg_ordered %>%
  filter(GAME_DATE_EST >= "2017-10-17" & GAME_DATE_EST <= "2018-4-11") %>%
  group_by(HOME_TEAM_ID) %>%
  summarise(GAME_DATE_EST = GAME_DATE_EST,
            GAME_ID = GAME_ID,
            PTS_home = PTS_home,
            games_each = n(),
            PTS_home_avg = sum(PTS_home)/games_each,
            OREB_home_avg = sum(OREB_home)/games_each,
            DREB_home_avg = sum(DREB_home)/games_each,
            FGM_home_avg = sum(FGM_home)/games_each,
            FGA_home_avg = sum(FGA_home)/games_each,
            FG3M_home_avg = sum(FG3M_home)/games_each,
            FG3A_home_avg = sum(FG3A_home)/games_each,
            FTM_home_avg = sum(FTM_home)/games_each,
            FTA_home_avg = sum(FTA_home)/games_each,
            AST_home_avg = sum(AST_home)/games_each,
            STL_home_avg = sum(STL_home)/games_each,
            BLK_home_avg = sum(BLK_home)/games_each,
            TO_home_avg = sum(TO_home)/games_each,
            PF_home_avg = sum(PF_home)/games_each,
            PM_home_avg = sum(PM_home)/games_each,
            PTS_home_against_avg = sum(PTS_away)/games_each,
            OREB_home_against_avg = sum(OREB_away)/games_each,
            DREB_home_against_avg = sum(DREB_away)/games_each,
            FGM_home_against_avg = sum(FGM_away)/games_each,
            FGA_home_against_avg = sum(FGA_away)/games_each,
            FG3M_home_against_avg = sum(FG3M_away)/games_each,
            FG3A_home_against_avg = sum(FG3A_away)/games_each,
            FTM_home_against_avg = sum(FTM_away)/games_each,
            FTA_home_against_avg = sum(FTA_away)/games_each,
            AST_home_against_avg = sum(AST_away)/games_each,
            STL_home_against_avg = sum(STL_away)/games_each,
            BLK_home_against_avg = sum(BLK_away)/games_each,
            TO_home_against_avg = sum(TO_away)/games_each,
            PF_home_against_avg = sum(PF_away)/games_each,
            PM_home_against_avg = sum(PM_away)/games_each)

head(season_level_avg_2017, 30)
```

```{r}
season_level_avg_2018 <- season_avg_ordered %>%
  filter(GAME_DATE_EST >= "2018-10-16" & GAME_DATE_EST <= "2019-4-10") %>%
  group_by(HOME_TEAM_ID) %>%
  summarise(GAME_DATE_EST = GAME_DATE_EST,
            GAME_ID = GAME_ID,
            PTS_home = PTS_home,
            games_each = n(),
            PTS_home_avg = sum(PTS_home)/games_each,
            OREB_home_avg = sum(OREB_home)/games_each,
            DREB_home_avg = sum(DREB_home)/games_each,
            FGM_home_avg = sum(FGM_home)/games_each,
            FGA_home_avg = sum(FGA_home)/games_each,
            FG3M_home_avg = sum(FG3M_home)/games_each,
            FG3A_home_avg = sum(FG3A_home)/games_each,
            FTM_home_avg = sum(FTM_home)/games_each,
            FTA_home_avg = sum(FTA_home)/games_each,
            AST_home_avg = sum(AST_home)/games_each,
            STL_home_avg = sum(STL_home)/games_each,
            BLK_home_avg = sum(BLK_home)/games_each,
            TO_home_avg = sum(TO_home)/games_each,
            PF_home_avg = sum(PF_home)/games_each,
            PM_home_avg = sum(PM_home)/games_each,
            PTS_home_against_avg = sum(PTS_away)/games_each,
            OREB_home_against_avg = sum(OREB_away)/games_each,
            DREB_home_against_avg = sum(DREB_away)/games_each,
            FGM_home_against_avg = sum(FGM_away)/games_each,
            FGA_home_against_avg = sum(FGA_away)/games_each,
            FG3M_home_against_avg = sum(FG3M_away)/games_each,
            FG3A_home_against_avg = sum(FG3A_away)/games_each,
            FTM_home_against_avg = sum(FTM_away)/games_each,
            FTA_home_against_avg = sum(FTA_away)/games_each,
            AST_home_against_avg = sum(AST_away)/games_each,
            STL_home_against_avg = sum(STL_away)/games_each,
            BLK_home_against_avg = sum(BLK_away)/games_each,
            TO_home_against_avg = sum(TO_away)/games_each,
            PF_home_against_avg = sum(PF_away)/games_each,
            PM_home_against_avg = sum(PM_away)/games_each)

head(season_level_avg_2018, 30)
```

```{r}
season_level_avg_2019 <- season_avg_ordered %>%
  filter(GAME_DATE_EST >= "2018-10-22" & GAME_DATE_EST <= "2019-8-14") %>%
  group_by(HOME_TEAM_ID) %>%
  summarise(GAME_DATE_EST = GAME_DATE_EST,
            GAME_ID = GAME_ID,
            PTS_home = PTS_home,
            games_each = n(),
            PTS_home_avg = sum(PTS_home)/games_each,
            OREB_home_avg = sum(OREB_home)/games_each,
            DREB_home_avg = sum(DREB_home)/games_each,
            FGM_home_avg = sum(FGM_home)/games_each,
            FGA_home_avg = sum(FGA_home)/games_each,
            FG3M_home_avg = sum(FG3M_home)/games_each,
            FG3A_home_avg = sum(FG3A_home)/games_each,
            FTM_home_avg = sum(FTM_home)/games_each,
            FTA_home_avg = sum(FTA_home)/games_each,
            AST_home_avg = sum(AST_home)/games_each,
            STL_home_avg = sum(STL_home)/games_each,
            BLK_home_avg = sum(BLK_home)/games_each,
            TO_home_avg = sum(TO_home)/games_each,
            PF_home_avg = sum(PF_home)/games_each,
            PM_home_avg = sum(PM_home)/games_each,
            PTS_home_against_avg = sum(PTS_away)/games_each,
            OREB_home_against_avg = sum(OREB_away)/games_each,
            DREB_home_against_avg = sum(DREB_away)/games_each,
            FGM_home_against_avg = sum(FGM_away)/games_each,
            FGA_home_against_avg = sum(FGA_away)/games_each,
            FG3M_home_against_avg = sum(FG3M_away)/games_each,
            FG3A_home_against_avg = sum(FG3A_away)/games_each,
            FTM_home_against_avg = sum(FTM_away)/games_each,
            FTA_home_against_avg = sum(FTA_away)/games_each,
            AST_home_against_avg = sum(AST_away)/games_each,
            STL_home_against_avg = sum(STL_away)/games_each,
            BLK_home_against_avg = sum(BLK_away)/games_each,
            TO_home_against_avg = sum(TO_away)/games_each,
            PF_home_against_avg = sum(PF_away)/games_each,
            PM_home_against_avg = sum(PM_away)/games_each)

head(season_level_avg_2019, 30)
```

# Find away team season-level averages for 2014-2019
```{r}
season_level_avg_away_2014 <- season_avg_ordered %>%
  filter(GAME_DATE_EST <= "2015-4-15") %>%  
  group_by(VISITOR_TEAM_ID) %>%
  summarise(GAME_DATE_EST = GAME_DATE_EST,
            GAME_ID = GAME_ID,
            PTS_away = PTS_away,
            games_each = n(),
            PTS_away_avg = sum(PTS_away)/games_each,
            OREB_away_avg = sum(OREB_away)/games_each,
            DREB_away_avg = sum(DREB_away)/games_each,
            FGM_away_avg = sum(FGM_away)/games_each,
            FGA_away_avg = sum(FGA_away)/games_each,
            FG3M_away_avg = sum(FG3M_away)/games_each,
            FG3A_away_avg = sum(FG3A_away)/games_each,
            FTM_away_avg = sum(FTM_away)/games_each,
            FTA_away_avg = sum(FTA_away)/games_each,
            AST_away_avg = sum(AST_away)/games_each,
            STL_away_avg = sum(STL_away)/games_each,
            BLK_away_avg = sum(BLK_away)/games_each,
            TO_away_avg = sum(TO_away)/games_each,
            PF_away_avg = sum(PF_away)/games_each,
            PM_away_avg = sum(PM_away)/games_each,
            PTS_away_against_avg = sum(PTS_home)/games_each,
            OREB_away_against_avg = sum(OREB_home)/games_each,
            DREB_away_against_avg = sum(DREB_home)/games_each,
            FGM_away_against_avg = sum(FGM_home)/games_each,
            FGA_away_against_avg = sum(FGA_home)/games_each,
            FG3M_away_against_avg = sum(FG3M_home)/games_each,
            FG3A_away_against_avg = sum(FG3A_home)/games_each,
            FTM_away_against_avg = sum(FTM_home)/games_each,
            FTA_away_against_avg = sum(FTA_home)/games_each,
            AST_away_against_avg = sum(AST_home)/games_each,
            STL_away_against_avg = sum(STL_home)/games_each,
            BLK_away_against_avg = sum(BLK_home)/games_each,
            TO_away_against_avg = sum(TO_home)/games_each,
            PF_away_against_avg = sum(PF_home)/games_each,
            PM_away_against_avg = sum(PM_home)/games_each)


head(season_level_avg_away_2014)
```

```{r}
season_level_avg_away_2015 <- season_avg_ordered %>%
  filter(GAME_DATE_EST >= "2015-10-27" & GAME_DATE_EST <= "2016-4-13") %>%  
  group_by(VISITOR_TEAM_ID) %>%
  summarise(GAME_DATE_EST = GAME_DATE_EST,
            GAME_ID = GAME_ID,
            PTS_away = PTS_away,
            games_each = n(),
            PTS_away_avg = sum(PTS_away)/games_each,
            OREB_away_avg = sum(OREB_away)/games_each,
            DREB_away_avg = sum(DREB_away)/games_each,
            FGM_away_avg = sum(FGM_away)/games_each,
            FGA_away_avg = sum(FGA_away)/games_each,
            FG3M_away_avg = sum(FG3M_away)/games_each,
            FG3A_away_avg = sum(FG3A_away)/games_each,
            FTM_away_avg = sum(FTM_away)/games_each,
            FTA_away_avg = sum(FTA_away)/games_each,
            AST_away_avg = sum(AST_away)/games_each,
            STL_away_avg = sum(STL_away)/games_each,
            BLK_away_avg = sum(BLK_away)/games_each,
            TO_away_avg = sum(TO_away)/games_each,
            PF_away_avg = sum(PF_away)/games_each,
            PM_away_avg = sum(PM_away)/games_each,
            PTS_away_against_avg = sum(PTS_home)/games_each,
            OREB_away_against_avg = sum(OREB_home)/games_each,
            DREB_away_against_avg = sum(DREB_home)/games_each,
            FGM_away_against_avg = sum(FGM_home)/games_each,
            FGA_away_against_avg = sum(FGA_home)/games_each,
            FG3M_away_against_avg = sum(FG3M_home)/games_each,
            FG3A_away_against_avg = sum(FG3A_home)/games_each,
            FTM_away_against_avg = sum(FTM_home)/games_each,
            FTA_away_against_avg = sum(FTA_home)/games_each,
            AST_away_against_avg = sum(AST_home)/games_each,
            STL_away_against_avg = sum(STL_home)/games_each,
            BLK_away_against_avg = sum(BLK_home)/games_each,
            TO_away_against_avg = sum(TO_home)/games_each,
            PF_away_against_avg = sum(PF_home)/games_each,
            PM_away_against_avg = sum(PM_home)/games_each)

head(season_level_avg_away_2015)
```

```{r}
season_level_avg_away_2016 <- season_avg_ordered %>%
  filter(GAME_DATE_EST >= "2016-10-25" & GAME_DATE_EST <= "2017-4-13") %>%  
  group_by(VISITOR_TEAM_ID) %>%
  summarise(GAME_DATE_EST = GAME_DATE_EST,
            GAME_ID = GAME_ID,
            PTS_away = PTS_away,
            games_each = n(),
            PTS_away_avg = sum(PTS_away)/games_each,
            OREB_away_avg = sum(OREB_away)/games_each,
            DREB_away_avg = sum(DREB_away)/games_each,
            FGM_away_avg = sum(FGM_away)/games_each,
            FGA_away_avg = sum(FGA_away)/games_each,
            FG3M_away_avg = sum(FG3M_away)/games_each,
            FG3A_away_avg = sum(FG3A_away)/games_each,
            FTM_away_avg = sum(FTM_away)/games_each,
            FTA_away_avg = sum(FTA_away)/games_each,
            AST_away_avg = sum(AST_away)/games_each,
            STL_away_avg = sum(STL_away)/games_each,
            BLK_away_avg = sum(BLK_away)/games_each,
            TO_away_avg = sum(TO_away)/games_each,
            PF_away_avg = sum(PF_away)/games_each,
            PM_away_avg = sum(PM_away)/games_each,
            PTS_away_against_avg = sum(PTS_home)/games_each,
            OREB_away_against_avg = sum(OREB_home)/games_each,
            DREB_away_against_avg = sum(DREB_home)/games_each,
            FGM_away_against_avg = sum(FGM_home)/games_each,
            FGA_away_against_avg = sum(FGA_home)/games_each,
            FG3M_away_against_avg = sum(FG3M_home)/games_each,
            FG3A_away_against_avg = sum(FG3A_home)/games_each,
            FTM_away_against_avg = sum(FTM_home)/games_each,
            FTA_away_against_avg = sum(FTA_home)/games_each,
            AST_away_against_avg = sum(AST_home)/games_each,
            STL_away_against_avg = sum(STL_home)/games_each,
            BLK_away_against_avg = sum(BLK_home)/games_each,
            TO_away_against_avg = sum(TO_home)/games_each,
            PF_away_against_avg = sum(PF_home)/games_each,
            PM_away_against_avg = sum(PM_home)/games_each)

head(season_level_avg_away_2016)
```

```{r}
season_level_avg_away_2017 <- season_avg_ordered %>%
  filter(GAME_DATE_EST >= "2017-10-17" & GAME_DATE_EST <= "2018-4-11") %>%  
  group_by(VISITOR_TEAM_ID) %>%
  summarise(GAME_DATE_EST = GAME_DATE_EST,
            GAME_ID = GAME_ID,
            PTS_away = PTS_away,
            games_each = n(),
            PTS_away_avg = sum(PTS_away)/games_each,
            OREB_away_avg = sum(OREB_away)/games_each,
            DREB_away_avg = sum(DREB_away)/games_each,
            FGM_away_avg = sum(FGM_away)/games_each,
            FGA_away_avg = sum(FGA_away)/games_each,
            FG3M_away_avg = sum(FG3M_away)/games_each,
            FG3A_away_avg = sum(FG3A_away)/games_each,
            FTM_away_avg = sum(FTM_away)/games_each,
            FTA_away_avg = sum(FTA_away)/games_each,
            AST_away_avg = sum(AST_away)/games_each,
            STL_away_avg = sum(STL_away)/games_each,
            BLK_away_avg = sum(BLK_away)/games_each,
            TO_away_avg = sum(TO_away)/games_each,
            PF_away_avg = sum(PF_away)/games_each,
            PM_away_avg = sum(PM_away)/games_each,
            PTS_away_against_avg = sum(PTS_home)/games_each,
            OREB_away_against_avg = sum(OREB_home)/games_each,
            DREB_away_against_avg = sum(DREB_home)/games_each,
            FGM_away_against_avg = sum(FGM_home)/games_each,
            FGA_away_against_avg = sum(FGA_home)/games_each,
            FG3M_away_against_avg = sum(FG3M_home)/games_each,
            FG3A_away_against_avg = sum(FG3A_home)/games_each,
            FTM_away_against_avg = sum(FTM_home)/games_each,
            FTA_away_against_avg = sum(FTA_home)/games_each,
            AST_away_against_avg = sum(AST_home)/games_each,
            STL_away_against_avg = sum(STL_home)/games_each,
            BLK_away_against_avg = sum(BLK_home)/games_each,
            TO_away_against_avg = sum(TO_home)/games_each,
            PF_away_against_avg = sum(PF_home)/games_each,
            PM_away_against_avg = sum(PM_home)/games_each)

head(season_level_avg_away_2017)
```

```{r}
season_level_avg_away_2018 <- season_avg_ordered %>%
  filter(GAME_DATE_EST >= "2018-10-16" & GAME_DATE_EST <= "2019-4-10") %>%  
  group_by(VISITOR_TEAM_ID) %>%
  summarise(GAME_DATE_EST = GAME_DATE_EST,
            GAME_ID = GAME_ID,
            PTS_away = PTS_away,
            games_each = n(),
            PTS_away_avg = sum(PTS_away)/games_each,
            OREB_away_avg = sum(OREB_away)/games_each,
            DREB_away_avg = sum(DREB_away)/games_each,
            FGM_away_avg = sum(FGM_away)/games_each,
            FGA_away_avg = sum(FGA_away)/games_each,
            FG3M_away_avg = sum(FG3M_away)/games_each,
            FG3A_away_avg = sum(FG3A_away)/games_each,
            FTM_away_avg = sum(FTM_away)/games_each,
            FTA_away_avg = sum(FTA_away)/games_each,
            AST_away_avg = sum(AST_away)/games_each,
            STL_away_avg = sum(STL_away)/games_each,
            BLK_away_avg = sum(BLK_away)/games_each,
            TO_away_avg = sum(TO_away)/games_each,
            PF_away_avg = sum(PF_away)/games_each,
            PM_away_avg = sum(PM_away)/games_each,
            PTS_away_against_avg = sum(PTS_home)/games_each,
            OREB_away_against_avg = sum(OREB_home)/games_each,
            DREB_away_against_avg = sum(DREB_home)/games_each,
            FGM_away_against_avg = sum(FGM_home)/games_each,
            FGA_away_against_avg = sum(FGA_home)/games_each,
            FG3M_away_against_avg = sum(FG3M_home)/games_each,
            FG3A_away_against_avg = sum(FG3A_home)/games_each,
            FTM_away_against_avg = sum(FTM_home)/games_each,
            FTA_away_against_avg = sum(FTA_home)/games_each,
            AST_away_against_avg = sum(AST_home)/games_each,
            STL_away_against_avg = sum(STL_home)/games_each,
            BLK_away_against_avg = sum(BLK_home)/games_each,
            TO_away_against_avg = sum(TO_home)/games_each,
            PF_away_against_avg = sum(PF_home)/games_each,
            PM_away_against_avg = sum(PM_home)/games_each)

head(season_level_avg_away_2018)
```

```{r}
season_level_avg_away_2019 <- season_avg_ordered %>%
  filter(GAME_DATE_EST >= "2018-10-22" & GAME_DATE_EST <= "2019-8-14") %>%  
  group_by(VISITOR_TEAM_ID) %>%
  summarise(GAME_DATE_EST = GAME_DATE_EST,
            GAME_ID = GAME_ID,
            PTS_away = PTS_away,
            games_each = n(),
            PTS_away_avg = sum(PTS_away)/games_each,
            OREB_away_avg = sum(OREB_away)/games_each,
            DREB_away_avg = sum(DREB_away)/games_each,
            FGM_away_avg = sum(FGM_away)/games_each,
            FGA_away_avg = sum(FGA_away)/games_each,
            FG3M_away_avg = sum(FG3M_away)/games_each,
            FG3A_away_avg = sum(FG3A_away)/games_each,
            FTM_away_avg = sum(FTM_away)/games_each,
            FTA_away_avg = sum(FTA_away)/games_each,
            AST_away_avg = sum(AST_away)/games_each,
            STL_away_avg = sum(STL_away)/games_each,
            BLK_away_avg = sum(BLK_away)/games_each,
            TO_away_avg = sum(TO_away)/games_each,
            PF_away_avg = sum(PF_away)/games_each,
            PM_away_avg = sum(PM_away)/games_each,
            PTS_away_against_avg = sum(PTS_home)/games_each,
            OREB_away_against_avg = sum(OREB_home)/games_each,
            DREB_away_against_avg = sum(DREB_home)/games_each,
            FGM_away_against_avg = sum(FGM_home)/games_each,
            FGA_away_against_avg = sum(FGA_home)/games_each,
            FG3M_away_against_avg = sum(FG3M_home)/games_each,
            FG3A_away_against_avg = sum(FG3A_home)/games_each,
            FTM_away_against_avg = sum(FTM_home)/games_each,
            FTA_away_against_avg = sum(FTA_home)/games_each,
            AST_away_against_avg = sum(AST_home)/games_each,
            STL_away_against_avg = sum(STL_home)/games_each,
            BLK_away_against_avg = sum(BLK_home)/games_each,
            TO_away_against_avg = sum(TO_home)/games_each,
            PF_away_against_avg = sum(PF_home)/games_each,
            PM_away_against_avg = sum(PM_home)/games_each)

head(season_level_avg_away_2019)
```

# Joining averages back with game matchup dataset
```{r}
avg_joined_2014 <- season_level_avg_2014 %>%
  left_join(season_level_avg_away_2014, by=c("GAME_ID"))

avg_joined_2015 <- season_level_avg_2015 %>%
  left_join(season_level_avg_away_2015, by=c("GAME_ID"))

avg_joined_2016 <- season_level_avg_2016 %>%
  left_join(season_level_avg_away_2016, by=c("GAME_ID"))

avg_joined_2017 <- season_level_avg_2017 %>%
  left_join(season_level_avg_away_2017, by=c("GAME_ID"))

avg_joined_2018 <- season_level_avg_2018 %>%
  left_join(season_level_avg_away_2018, by=c("GAME_ID"))

avg_joined_2019 <- season_level_avg_2019 %>%
  left_join(season_level_avg_away_2019, by=c("GAME_ID"))

head(avg_joined_2014)
head(avg_joined_2015)
head(avg_joined_2016)
head(avg_joined_2017)
head(avg_joined_2018)
head(avg_joined_2019)
```

# Compile each year into same dataset
```{r}
avg_and_game <- rbind(avg_joined_2014, avg_joined_2015, avg_joined_2016, avg_joined_2017, avg_joined_2018, avg_joined_2019)

head(avg_and_game)
```


# Create dataset of only season-level averages

```{r}
season_level_only <- avg_and_game %>%
  select(-games_each.x, -games_each.y, -GAME_DATE_EST.y) %>%
  rename("GAME_DATE_EST" = GAME_DATE_EST.x)

head(season_level_only)
```

```{r}
head(season_avg_ordered)
```

```{r}
joining_table <- season_avg_ordered[, c("GAME_DATE_EST", "GAME_ID", "Home.Team", "Away.Team", "Spread", "Total", "OREB")]
head(joining_table)
```



```{r}
season_level_only_1 <-  merge(season_level_only, joining_table, by = c("GAME_ID"))

head(season_level_only_1)
```


# Reordering variables
```{r}
season_level_only_final <- season_level_only_1[, c("GAME_DATE_EST.x", "Home.Team", "Away.Team", "Spread", "Total", "OREB", "PTS_home", "PTS_away", "PTS_home_avg", "PTS_home_against_avg", "PTS_away_avg", "PTS_away_against_avg", "OREB_home_avg", "OREB_home_against_avg", "OREB_away_avg", "OREB_away_against_avg", "DREB_home_avg", "DREB_home_against_avg", "DREB_away_avg", "DREB_away_against_avg", "FGM_home_avg", "FGM_home_against_avg", "FGM_away_avg", "FGM_away_against_avg", "FGA_home_avg", "FGA_home_against_avg", "FGA_away_avg", "FGA_away_against_avg", "FG3M_home_avg", "FG3M_home_against_avg", "FG3M_away_avg", "FG3M_away_against_avg", "FG3A_home_avg", "FG3A_home_against_avg", "FG3A_away_avg", "FG3A_away_against_avg", "FTM_home_avg", "FTM_home_against_avg", "FTM_away_avg", "FTM_away_against_avg", "FTA_home_avg", "FTA_home_against_avg", "FTA_away_avg", "FTA_away_against_avg", "AST_home_avg", "AST_home_against_avg", "AST_away_avg", "AST_away_against_avg", "STL_home_avg", "STL_home_against_avg", "STL_away_avg", "STL_away_against_avg", "BLK_home_avg", "BLK_home_against_avg", "BLK_away_avg", "BLK_away_against_avg", "TO_home_avg", "TO_home_against_avg", "TO_away_avg", "TO_away_against_avg", "PF_home_avg", "PF_home_against_avg", "PF_away_avg", "PF_away_against_avg", "PM_home_avg", "PM_home_against_avg", "PM_away_avg", "PM_away_against_avg") ]

season_level_stats <- season_level_only_final %>%
  rename("GAME_DATE_EST"= GAME_DATE_EST.x)

head(season_level_stats,100)
```


# Write CSV
```{r}
write.csv(season_level_stats, 'season_level_stats.csv')
write.csv(season_level_stats, 'season_level_with_points.csv')
```

```{r}
write.csv(season_level_stats, 'season_level_for_against.csv')
```









Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.