---
title: "Spread Regression"
output: html_notebook
---

```{r}
library(MASS)
```


# Importing game data
```{r}
games <-  read.csv("https://raw.githubusercontent.com/mattymo18/STOR-538-Project2-2021/master/Source-Data/games.csv")
game_details <- read.csv("https://raw.githubusercontent.com/mattymo18/STOR-538-Project2-2021/master/Source-Data/games_details.csv")
teams <- read.csv("https://raw.githubusercontent.com/mattymo18/STOR-538-Project2-2021/master/Source-Data/teams.csv")
all_game_data <- read.csv("https://raw.githubusercontent.com/john-lickteig/playoff2/master/all_game_data.csv")
game_details_cleaned <- read.csv("https://raw.githubusercontent.com/john-lickteig/playoff2/master/game_details_cleaned.csv")
games_cleaned <- read.csv("https://raw.githubusercontent.com/john-lickteig/playoff2/master/games_cleaned.csv")
final_game_data_with_rolling_avg <- read.csv("https://raw.githubusercontent.com/john-lickteig/playoff2/master/final_game_data_with_rolling_avg.csv")
```

#Define Stepwise regression model

First we define a model with all predictors, removing the ones at the beginning and PTS_home and PTS_away because they are add up to the spread.
```{r}
all_predictors = lm(Spread ~ . - X - GAME_DATE_EST - Home.Team - Away.Team - PTS_home - PTS_away, data=final_game_data_with_rolling_avg)
```



Next, we use the function stepAIC and use forward selection.
```{r}
forward = stepAIC(all_predictors, direction='forward', trace=FALSE)

summary(forward)
```
We find the significant predictors to be Total, FGM_home, FGM_away, FG3M_home, FG3M_away, FTM_home, FTM_away, PM_home, and PM_away.

# Define a new model using the results from forward selection
```{r}
forward_model = lm(Spread ~ Total + FGM_home + FGM_away + FG3M_home + FG3M_away + FTM_home + FTM_away + PM_home + PM_away, data=final_game_data_with_rolling_avg)
```



Then we use backward selection.
```{r}
backward = stepAIC(all_predictors, direction='backward', trace=FALSE)

summary(backward)
```
We find the significant predictors here are also Total, FGM_home, FGM_away, FG3M_home, FG3M_away, FTM_home, FTM_away, PM_home, and PM_away.



# Define a new model using the results from backward selection
```{r}
backward_model = lm(Spread ~ Total + FGM_home + FGM_away + FG3M_home + FG3M_away + FTM_home + FTM_away + PM_home + PM_away, data=final_game_data_with_rolling_avg)
```

Each selection process found that the same predictors were significant.





.

